\plaintitle

### 2017-02-24 paper/paper.tex
Paper: Better captions
Paper: Discussion and future research draft

The current study examines to what extent physical effort can enhance expressivity in playing electronic instruments.
To address this question I use the Schleikess interface, a digital controller for interactive performance that requires physical effort and large body movements from the performer.
Figure \ref{fig:schleikess} shows a schematic diagram of the interface.
It is composed of two elastic bands, each of them is connected with one side, through a load sensor, to a belt loop of the performer.
The load sensor measures the tension of the elastic band as it stretches and transmit this information to the computer to translate it to sound.
The performer holds the other side of the elastic bands and can stretch them to play.

\begin{figure}
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/schleikess}
  \caption{A diagram of the Schleikess interface. In blue are the two elastic bands. Stretching the left raises the pitch and stretching the right raises the amplitude of a sawtooth tone generator.}~\label{fig:schleikess}
\end{figure}

@@

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\subsection{Future research}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

>>>>

The current study examines to what extent physical effort can enhance expressivity in playing electronic instruments.
To address this question I use the Schleikess interface, a digital controller for interactive performance that requires physical effort and large body movements from the performer.
Figure \ref{fig:schleikess} shows the interface design.
It is composed of two elastic bands, each of them is connected with one side, through a load sensor, to a belt loop of the performer.
The load sensor measures the tension of the elastic band as it stretches and transmit this information to the computer to translate it to sound.
The performer holds the other side of the elastic bands and can stretch them to play.

\begin{figure}
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/schleikess}
  \caption{The design of the Schleikess interface. In blue are the two elastic bands. Stretching the left raises the pitch and stretching the right raises the amplitude of a sawtooth tone generator.}~\label{fig:schleikess}
\end{figure}

@@

The preliminary results presented in this study only partially show the effect of physical effort on expressivity in electronic instruments performance.
The most significant pattern emerged from the experiment is the common trade-off between controllability and range of possibilities, especially in the pitch dimension.
Despite the fact that this trade-off was mediated by the sensitivity of the interface to force, the required force by itself wasn't the main consideration of the participants.
In other words, if this trade-off was mediated, by design, by another aspect of the interface, I suspect that the participants answers were similar.

On the other hand, as well as the required physical effort, the Schleikess also provides physical feedback for the player.
This feedback cannot be separated from the physicality of the interface, and it is an integral part in all of the interfaces of this kind.
I believe that this feedback provided greater controllability and at the same time enhanced the openness of the participants to elements of randomness in their play.
This claims, however, weren't assessed directly by the current study.

\section{Future research}

The current study highlighted the importance of controllability and the decisions users made when facing trade-offs between controllability and range of possibilities.
To address the question of expressivity and its relationship to physical effort a more comprehensive study should be conducted.
Future research in the area should explore both the open question of whether and how physical effort can enhance expressivity in playing electronic instruments.
In addition, it should compare the results of performance expressivity from the point of view of the performer to the listener experience of the performance.
Although this type of investigation was beyond the scope of the current research it is crucial in order to understand these relationships better.
